---
title: "Prediction Assignment Writeup"
author: "Tanya Bisht"
date: "4/15/2020"
output: html_document
---

I have tried to generate correct answers for each of the 20 test data cases of this assignment using three modeles. The model with the highest accuracy is used to predict the quiz results.  

```{r}
library(knitr)
library(caret)
library(rpart)
library(randomForest)
library(psych)
library(e1071)
library(rpart)
library(rpart.plot)
set.seed(301)
```

```{r}
TrainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
TestUrl  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
TrainFile<-"pml-training.csv"
TestFile<-"pml-testing.csv"

if(!file.exists(TrainFile))
{
    download.file(TrainUrl,destfile = TrainFile)
}
training <- read.csv(TrainFile)
if(!file.exists(TestFile))
{
    download.file(TestUrl,destfile = TestFile)
}
testing  <- read.csv(TestFile)
```

#Create a partition using caret with the training dataset on 70,30 ratio
```{r}
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)

TrainSet <- training[inTrain, ]

TestSet  <- training[-inTrain, ]
dim(TrainSet)
dim(TestSet)
```

#Remove variables with Nearly Zero Variance
```{r}
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
dim(TestSet)

dim(TrainSet)
```

#Remove variables that are mostly NA
```{r}
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
dim(TestSet)
dim(TrainSet)
```

#Remove identification only variables (columns 1 to 5)
```{r}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

##Correction Analysis
```{r}
corMatrix <- cor(TrainSet[, -54])
corPlot(corMatrix)
```


##Prediction Model Building
#Random Forests
```{r}
set.seed(301)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf",
                          trControl=controlRF)
modFitRandForest$finalModel

predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest

plot(confMatRandForest$table, col = confMatRandForest$byClass, 
     main = paste("Random Forest - Accuracy =",
                  round(confMatRandForest$overall['Accuracy'], 4)))
```

#Decision Tree
```{r}
set.seed(301)
modFitDecTree <- rpart(classe ~ ., data=TrainSet, method="class")
rpart.plot(modFitDecTree)

predictDecTree <- predict(modFitDecTree, newdata=TestSet, type="class")
confMatDecTree <- confusionMatrix(predictDecTree, TestSet$classe)
confMatDecTree

plot(confMatDecTree$table, col = confMatDecTree$byClass, 
     main = paste("Decision Tree - Accuracy =",
                  round(confMatDecTree$overall['Accuracy'], 4)))
```

#Generalized Boosted Model (GBM)
```{r}
set.seed(301)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel

predictGBM <- predict(modFitGBM, newdata=TestSet)
confMatGBM <- confusionMatrix(predictGBM, TestSet$classe)
confMatGBM

plot(confMatGBM$table, col = confMatGBM$byClass, 
     main = paste("GBM - Accuracy =", round(confMatGBM$overall['Accuracy'], 4)))
```

##Applying Random Model to the Test Data
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```
